# Machine Learning - Barbell

### Import the data and packages 
assumes Caret package is already installed and data files are in current working directory

```{r}
#Import data file, converting all NA like values to NA
data <- read.csv("pml-training.csv", na.strings=c("NA", "N/A","-", "", "NAN", "Inf", "Div/0", "#DIV/0!"))

#Open Caret package
library(caret)
```

### Split into training and testing
```{r}
set.seed(1999)
train<-createDataPartition (data$classe, p=.6, list=FALSE)
training <- data[train,]
testing <- data[-train,]
```


### Select variables
Remove variabes that are inappropriate (subject name, date of event) and/or unuseful (variables with more than 50% NA values)
```{r}
#Remove username and timestamp columns
training <- training [,-(1:7)]
testing <- testing [,-(1:7)]

#Find and keep columns with less than 50% NA
na_50_less <- sapply(training, function(x){percent_na <- sum(is.na(x)) / length(x); percent_na<0.5})
training <- training [,na_50_less]
testing <- testing [,na_50_less]
```


### Preprocess data with PCA
```{r}
training_pca <- preProcess(training[,-ncol(training)], method="pca")

training_pca_pred <- predict(training_pca, training[,-ncol(training)])
testing_pca_pred <- predict(training_pca, testing[,-ncol(testing)])
```

### Train Random forest model
```{r, cache = TRUE}
train_control = trainControl(method = "cv", number=5)
set.seed(42)
model_rf <- train(training$classe~., method="rf", data=training_pca_pred, trControl=train_control)

# See results
model_rf$finalModel
```

NOTE: Cross-validation is performed automatically by the Random Forest training method above. 

The estimate out of bag error rate is XXXXX. Below we will test the model on the testing set to get a better ide of the out of bag error

### Predict on my testing set
```{r}
#predict on my testing set
pred_test <- predict(model_rf, testing_pca_pred)

#output confusion matrix to see how model fared
confusionMatrix(pred_test, testing$classe)
```

As you can see the accuracy is XXXX and so the out of bag error rate is XXXXX (1-XXXXX).

```{r, echo=FALSE}
#Import the 20 assignment test cases
assignment_20 <- read.csv("pml-testing.csv")

#Remove username and timestamp columns
assignment_20 <- assignment_20 [,-(1:7)]

#Keep columns with less than 50% NA
assignment_20 <- assignment_20 [,na_50_less]

#Preprocess data with PCA
assignment_20_pca_pred <- predict(training_pca, assignment_20[,-ncol(assignment_20)])

#Predict on 20 assignment test cases
pred_assignment_20 <- predict(model_rf, assignment_20_pca_pred)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(pred_assignment_20)
```

constrain the text of the writeup to < 2000 words and the number of figures to be less than 5
